### 2016-08-24 image_setup.tex
page 1 to 10 corrections

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
All other files used in this paper are available there unless specified
otherwise. To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator. To copy the commands from this procedure to the
testbed administrator terminal, we recommend to read this document
with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
render properly when pasted.

@@

In this code block, the \texttt{\$\{URL\}} and \texttt{\$\{IMAGE\}}
variables specify where the Linux bundle is located and
\texttt{\$\{WORKDIR\}} specifies a working directory where the Raspbian
Lite bundle will be downloaded and customized. Notice that even though we
use the \$ and \# signs in the shell, in general these signs will be
particular to the testbed administrator \ac{OS} shell. Next, we create
the working directory and go to it with the \texttt{cd} command. To
download the image, we utilize the \texttt{wget} command before unpacking
the \texttt{zip} file as follows:

@@

After Raspbian Lite has been unpacked, there should be an \texttt{.img}
file in the \texttt{\$\{WORKDIR\}} directory. Here, \texttt{fdisk} can
be used to display the content of the image. To do this and obtain
administrative information, we parse the arguments \texttt{-u sectors}
to display the sizes in sectors and \texttt{-l} to display the partitions
within the image. The \texttt{fdisk} command should output to the terminal
something similar to:

@@

The output provides relevant information about the image. The image is in
total 2709504 sectors (1.3 GiB) in size and contains two partitions. The
first partition starts at sector 8192 and the other partition starts at
sector 137216. The first partition type is FAT32 with a size of 63 MB
and the second partition is a Linux one with a size of 1.2 GB. This
indicates that the first partition is a boot partition and the second
one is a traditional Linux root filesystem, i.e. \texttt{/}.

\subsection{Image Resizing}
Given that we want to customize the files stored in the \ac{Raspi}s,
we need to resize the image file since 1.2 GB might not be enough to store
the root filesystem due to the total size of the additional packages. Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

@@

Now, in this command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, it is required to first remove the Linux partition
and then add it again with one GB more. To do this, we make use
again of \texttt{fdisk} in command mode to alter the partition table.
For this, we pass the arguments of the command mode of \texttt{fdisk}
through the \texttt{echo} command in Linux and the \texttt{|} operator
(pipe) as follows:

@@

\subsection{Block Devices Mounting}
For browsing and altering the files in the image, we mount the block
device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
to customize them. We proceed first to mount the block device partition that
contains the root filesystem and later the boot partition. This is done by
creating an empty directory that it is used as a mountpoint. We name it
\texttt{root} and create it in the working directory before mounting the
root filesystem onto the mountpoint. Thus, we mount the root filesystem
as follows:

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order are the scripts executed to configure all the
devices. Nevertheless, any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

\subsubsection{Image Default Configuration Scripts Download}
\label{sec:configuration_files_download}
In our case, we have our default configuration scripts stored in a
\texttt{rasp\_config.zip} file located in the same \ac{URL} of the
\ac{HTTP} server where the image was retrieved from, i.e. the one in
the environment variable \texttt{\$\{URL\}}. We first manually download
this compressed file with \texttt{wget} and extract it locally into our
Raspbian Lite image. These commands and the output of the last one are
shown as follows:

@@

\subsubsection{Device Hostnames}
Among the settings that we want to be different in the devices is
their hostname. The hostname helps the user to physically distinguish the
devices from each other. Thus, in our case we require the devices in our
testbed to have different hostnames. Hence, we define the hostnames based
on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his / her \ac{Raspi}s obtained previously in the format shown in the
configuration file. In the file format, on each line there is a \ac{MAC}
address and the corresponding hostname for the given \ac{Raspi}. For
simplicity in the example above, we just include three \ac{Raspi}s.
This list file will be employed by the
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
\ac{Bash} script to assign the hostname of each \ac{Raspi}.
The script content is the following:

@@

In our case as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files but this will be made in an
automated wat after booting up the system. This script contents are:

@@

The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
folder in the corresponding \ac{Raspi}. It also prints a warning in case of
errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
directory, while ignoring existing files and / or directories there.

@@

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
Therefore, we first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file as root. The resulting file should look like the following:

@@

If it is required to have more configuration scripts, adding them
in the \texttt{rc.local} file difficults its maintenance by the testbed
administrator since this needs to be both in the image and the
downloaded \texttt{rasp\_config.zip}. To avoid this problem, we include
the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/main} script which
calls all other configuration scripts (besides \texttt{update\_rasp\_config})
in sequential order. This script content is the following:

@@

In this way, the automation process is simplify since we only need to modify
the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:

@@

\texttt{chroot} is a method in Linux that modifies the apparent root
filesystem location from \texttt{/} to any other path. Consequently, in
our case, we can use the Raspbian Lite image root filesystem within the
testbed administrator Linux distribution. Then, QEMU allows the execution
of commands for the \ac{Raspi} image (\ac{ARM} instructions) through the ones
from the testbed administrator \ac{PC} architecture. Due to the \ac{ARM}
processor that the \ac{Raspi}s employ, t is required to install the QEMU related software first and verigy that QEMU is \ac{ARM} enabled. To do so,
run the following commands:

@@

Finally, to change root the following needs to be run:

@@

The Raspbian lite image should now be possible to use almost as if it had
been booted in a \ac{Raspi}. A major difference is that the testbed
administrator \ac{PC} is likely significantly faster than a \ac{Raspi}.
Hence, enabling updates, upgrades and installing new software packages
should be much faster than in a \ac{Raspi}. Still updating and upgrading
the packages for the \ac{Raspi} might take some amount of time. To get
the upgraded system packages, it is necessary to run:

>>>>

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{tunescode_webpage} acting as our \ac{HTTP} server.
All other files used in this paper are available there unless specified
otherwise. To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

@@

In this code block, the \texttt{\$\{URL\}} and \texttt{\$\{IMAGE\}}
variables specify where the Linux bundle is located and
\texttt{\$\{WORKDIR\}} specifies a working directory where the Raspbian
Lite bundle will be downloaded and customized. Notice that even though we
use the \$ and \# signs in the shell, in general these signs will be
particular to the testbed administrator \ac{OS} shell. Next, we create
the working directory and go to it with the change directory (\texttt{cd}) command. To
download the image, we utilize the \texttt{wget} command before unpacking
the \texttt{zip} file as follows:

@@

After Raspbian Lite has been unpacked, there should be an \texttt{.img}
file in the working directory \texttt{\$\{WORKDIR\}}. Here, \texttt{fdisk} can
be used to display the content of the image. To do this and obtain
administrative information, we parse the arguments \texttt{-u sectors}
to display the sizes in sectors and \texttt{-l} to display the partitions
within the image. The \texttt{fdisk} command should output to the terminal
something similar to:

@@

The output provides relevant information about the image. The image is in
total 2709504 sectors (1.3 GiB) in size and contains two partitions. The
first partition starts at sector 8192 and the other partition starts at
sector 137216. The first partition type is FAT32 with a size of 63 MB
and the second partition is of type Linux with a size of 1.2 GB. This
indicates that the first partition is a boot partition and the second
one is a traditional Linux filesystem. In this case, the root filesystem, i.e. \texttt{/}.

\subsection{Image Resizing}
Given that we want to customize the root filesystem in the \ac{Raspi}s,
we need to expand the image file since 1.2 GB might not be enough to store
existing root filesystem plus additional files and software packages.
%Given that we want to customize the files stored in the \ac{Raspi}s,
%we need to resize the image file since 1.2 GB might not be enough to store
%the root filesystem with the additional packages that will be installed in this procedure.
%due to the total size of the additional packages.
Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

@@

Now, in this command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, it is required to first remove the Linux partition
and then add it again with one GB more. To do this, we make use
of \texttt{fdisk} again in command mode to alter the partition table.
We pass the arguments of the command mode of \texttt{fdisk}
through the \texttt{echo} command in Linux and the \texttt{|} operator
(pipe) as follows:

@@

\subsection{Block Devices Mounting}
For browsing and altering the files in the image, we mount the block
device partitions into a particular path of our \texttt{\$\{WORKDIR\}} in order
to customize them. We proceed first to mount the block device partition that
contains the root filesystem and later the boot partition. This is done by
creating an empty directory that is used as a mountpoint. We name it
\texttt{root} and create it in the working directory before mounting the
root filesystem onto the mountpoint. Thus, we mount the root filesystem
as follows:

@@

\subsection{Image OS Files and Configuration Scripts Setup}
In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image. For this, we first
indicate how to obtain and put our configuration scripts in the image. Later,
we describe the tasks performed by these configuration scripts. Finally, we
indicate how and in which order the scripts are executed to configure all the
devices. Nevertheless, any testbed administrator might modify or include other
tasks besides the presented in this procedure according to his / her needs
as we will show.

\subsubsection{Image Default Configuration Scripts Download}
\label{sec:configuration_files_download}
In our case, we have our default configuration scripts stored in a
file \texttt{rasp\_config.zip} located in the same \ac{URL} of the
\ac{HTTP} server where the image was retrieved from, i.e. the one in
the environment variable \texttt{\$\{URL\}}. We first download
this compressed file with \texttt{wget} and extract it locally into our
Raspbian Lite image. These commands and the output of the last one are
shown as follows:

@@

\subsubsection{Device Hostnames}
Among the settings that we want to be different in the devices is
their hostname. The hostname helps the user to physically distinguish the
devices from each other. Thus, in our case we require the devices in our
testbed to have different hostnames. We define the hostnames based
on the \ac{MAC} addresses of the \ac{Raspi}s wired Ethernet interface.

@@

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his / her \ac{Raspi}s obtained previously in the format shown in the
configuration file. In that file format, on each line there is a \ac{MAC}
address and the corresponding hostname for the given \ac{Raspi}. For
simplicity in the example above, we just include four \ac{Raspi}s.
This file will be employed by the
\texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}
\ac{Bash} script to assign the hostname of each \ac{Raspi}.
The script content is the following:

@@

In our case as the testbed administrator for presenting the procedure, we want
to fetch all our configuration scripts in
\texttt{\$\{URL\%/*\}/raspi\_config.zip}. Thus, the update script
\textit{automatically} downloads all the required configuration files in
\texttt{rasp\_config.zip} file from a remote location. This is the same that
we manually did earlier to get our files but this will be made in an
automated way after booting up the system. This script content is:

@@

The update script: (3-4) checks for the \ac{URL} and \texttt{.zip} file that
should be downloaded, (7-10) downloads the configuration files to \texttt{/tmp}
folder in the corresponding \ac{Raspi}. It also prints a warning in case of
errors and (13) unzips the files to \texttt{/home/pi/}, the \ac{Raspi} home
directory. Existing files and directories are simply overwritten.
%while ignoring existing files and / or directories there.

@@

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
Therefore, we first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file with root permissions. The resulting file should look like the following:

@@

If it is required to have more configuration scripts, adding them
in the \texttt{rc.local} file difficults its maintenance by the testbed
administrator since this needs to be both in the image and the
downloaded \texttt{rasp\_config.zip}. To avoid this problem, we include
the \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/main} script which
calls all other configuration scripts (besides \texttt{update\_rasp\_config})
in a sequential order. This script content is:

@@

In this way, the automation process is simplified since we only need to modify
the image \texttt{\$\{ROOTDIR\}/etc/rc.local} file once to execute this script:

@@

\texttt{chroot} is a method in Linux that modifies the apparent root
filesystem location from \texttt{/} to any other path. Consequently, in
our case, we can use the Raspbian Lite image root filesystem within the
testbed administrator Linux distribution. Then, QEMU allows the execution
of commands for the \ac{Raspi} image (\ac{ARM} instructions) through the ones
from the testbed administrator \ac{PC} architecture. Due to the \ac{ARM}
processor that the \ac{Raspi}s employ, it is required to install the QEMU related software first and verify that QEMU is \ac{ARM} enabled. To do so,
run the following commands:

@@

%Finally, to change root the following needs to be run:
Finally, run the following command to change root:

@@

The Raspbian lite image should now be possible to use almost as if it had
been booted in a \ac{Raspi}. A major difference is that the testbed
administrator \ac{PC} is likely significantly faster than a \ac{Raspi}.
Hence, enabling updates, upgrades and installing new software packages
should be faster than in a \ac{Raspi}. Still updating and upgrading
the packages for the \ac{Raspi} might take some amount of time. To get
the upgraded system packages, it is necessary to run:

### 2016-08-24 introduction.tex
page 1 to 10 corrections

The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale. This will break new ground for new markets such as driverless cars, manufacturing, humanoid robots, smart grids, and many more. With 5G, the number of wireless devices will increase by 5$\times$ up to 50 billion devices \cite{cisco2011forecast}. It is generally accepted, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connect due to the expected energy consumption. Therefore the communication architecture in future communication scenario will be dominated by mesh technologies. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

In all these prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities. However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude. Moreover, despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce. Furthermore, deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently. The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages. Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintenable fashion.

The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area. This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from widely-used \ac{OS}. Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using Raspberry Pi devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo. Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with Raspberry Pi. Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system. In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity. In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step. Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed t simplify the execution of routinary and repetitive tasks. Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}. Conclusions and future work are reviewed in Section~\ref{sec:conclusions}. Finally, a set of alternative commands procedures, in case the ones presented in this work might not be executed, are discussed in the Appendix.

>>>>

The upcoming 5G technology is targeting the controlling and steering of the Internet of Things (IoT) in real-time on global scale. This will break new ground for new markets such as driverless vehicles, manufacturing, humanoid robots, smart grids, and many more. With 5G, the number of wireless devices will increase by five times up to 50 billion devices \cite{cisco2011forecast}. It is generally acceptable, that those devices will not be connected in the same cellular manner as the current 10 billion mobile devices are connected due to the expected energy consumption. Therefore the communication architecture in future communication scenario will be dominated by mesh technologies. Mesh technology has been known for sensor and ad hoc networks or mobile cloud scenarios, but the technical requirements on 5G mesh-based communication systems are dramatically increasing. Future mesh networks need to support high data rate, low latency, security, network availability and heterogeneous devices. In state of the art systems those requirements are traded-off with each other, but in the 5G context we cannot do this anymore.

@@

In all these prior works, the C++11 Kodo library \cite{kodo2011pedersen} was used as the common building block containing the basic \ac{RLNC} functionalities. However, most of the work was focusing on small mesh networks with a handful of communication nodes, though the expected scenarios are fairly beyond this order of magnitude. Moreover, despite this successful deployment in real systems, many of these protocols and contributions have been implemented in separate testbeds and the experiences are hard to reproduce. Furthermore, deploying a large-scale and configurable testbed for networking and storage can be challenging, not only due to the inherent costs of the hardware, but due to maintenance challenges and ability to replicate results consistently. The latter requires not only the devices to run the same \ac{OS}, but also have exactly the same configurations and software packages. Hence, there is a need in evaluating large scale network deployments of low cost devices in a quick, easy-to-deploy, reproducible and maintainable fashion.

The emergence of powerful and inexpensive single-board computers running full versions of a standard \ac{OS} from many possible opens new possibilities in this area. This not only allows to seamlessly develop implementations that are compatible with higher end devices and allows the reuse stable software from widely-used \ac{OS}. Given this set of specific needs, in this work we present the design, key step-by-step instructions and mechanisms to setup, configurate and maintain of an inexpensive testbed using Raspberry Pi devices for networking (wireless or wired) and storage applications including \ac{RLNC} functionalities into the testbed through Kodo. Nevertheless, the architecture itself is not bounded to the networking area and can be used for other applications that require replicable results with Raspberry Pi. Our work for the testbed procedure is organized as follows: Section~\ref{sec:overview} introduces the testbed system. In Section~\ref{sec:image_setup}, we provide details about the testbed setup, scripts, configuration files and connectivity. In Section~\ref{sec:overlay_fs}, we elaborate on the need and setup for an overlay filesystem for our testbed in order to have both persistent and non-persistent data on it as an optional step. Section~\ref{sec:tools} describes a set of automation and monitoring tools that can be included in the testbed to simplify the execution of routinary and repetitive tasks. Section~\ref{sec:cross_compilation} elaborates on the compilation of the Kodo library for the \ac{Raspi}. Conclusions and future work are reviewed in Section~\ref{sec:conclusions}. Finally, a set of alternative commands, in case the ones presented in this work might not be executed, are discussed in the Appendix.

### 2016-08-24 overview.tex
page 1 to 10 corrections

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a 8 GB memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and
professors to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. Nevertheless, the
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s are configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

>>>>

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a 8 GB memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and
employees to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. Nevertheless, the
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s are configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP} address. This
eliminates the demand for monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.
